---
title: "Survival Practical"
author: "Mohd Azmi"
date: "20/03/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Environment

## Packages

```{r}
library(pacman)
p_load(haven, tidyverse, summarytools, broom, survival, SurvRegCensCov, flexsurv)
```

## Dataset
```{r}
addicts <- read_dta("addicts.dta")
addicts1 <- addicts %>% mutate_if(is.labelled, list(~ as_factor(.)))
```

# Analysis

## Exploration
```{r}
addicts1
descr(addicts1)
freq(addicts1)
```

ID - patient ID
survt - time in day until patient was drop out of clinic or censored
status - 0 censored, 1 drop out
clinic - clinic 1 or clinic 2
prison - 0 no record, 1 hx of prison
dose - methadone dose (continous variable)

```{r}
ggplot(addicts1) + geom_histogram(aes(survt))
ggplot(addicts1, aes(survt)) +
  geom_histogram(aes(y = ..density..)) +
  stat_function(fun = dnorm, args = list(mean = mean(addicts1$survt),
                                         sd = sd(addicts1$survt)))
ggplot(addicts1, aes(sample = survt)) +
  geom_qq() + geom_qq_line()
```


## estimation of AFT models

### Exponential model

```{r}
surv_mode <- Surv(time = addicts1$survt, event = addicts1$status == 1)
exp_mod <- survreg(Surv(survt, status == 1) ~ prison + dose + clinic, data = addicts1, dist = "exponential")
cox_mod <- coxph(Surv(survt, status == 1) ~ prison + dose + clinic, data = addicts1)
summary(exp_mod)
summary(cox_mod)
```

```{r}
tidy(exp_mod, conf.int = T) %>% 
  mutate(TR = exp(estimate),
         llci = exp(conf.low),
         ulci = exp(conf.high),
         HR = exp(-estimate)) %>% 
  select(term, estimate, TR, llci, ulci, p.value, HR)
```

estimated log time to drop out of the clinic in subject with past prison history (as compared to those without prison history) = `r round(coef(exp_mod)[2],3)`

AF or TR = `r round(exp(coef(exp_mod))[2],3)`

HR = `r round(exp(-coef(exp_mod))[2],3)`

### Weibull model

#### Weibull Model Option 1

```{r}
wei_mod <- survreg(Surv(survt, status == 1) ~ prison + dose + clinic, data = addicts1, dist = "weibull")
summary(wei_mod)
```

```{r}
tidy(wei_mod) %>% 
  mutate(TR = exp(estimate),
         llci = exp(conf.low),
         ulci = exp(conf.high),
         HR = exp(-estimate)) %>% 
  select(term, estimate, TR, llci, ulci, p.value, HR) #can't use this one because scale is 0.73, need to consider the scale

```

can't use this one because scale is 0.73, need to consider the scale


```{r}
ConvertWeibull(wei_mod, conf.level = 0.95) # to convert to weibull ph model
```



#### Weibull model - option 2

aft model

```{r}
wei_mod_o2_aft <- flexsurvreg(Surv(survt, status == 1) ~ prison + dose + clinic, data = addicts1, dist = "weibull")
wei_mod_o2_aft #same with survreg(Surv(survt, status == 1) ~ prison + dose + clinic, data = addicts1, dist = "weibull")
```

ph model

```{r}
wei_mod_o2_ph <- flexsurvreg(Surv(survt, status == 1) ~ prison + dose + clinic, data = addicts1, dist = "weibullPH")
wei_mod_o2_ph  #same with ConvertWeibull()
```


### Model Adequacy

```{r}
WeibullDiag(Surv(survt, status == 1) ~ prison, data = addicts1)
WeibullDiag(Surv(survt, status == 1) ~ clinic, data = addicts1)

```

## Prediction

```{r}
new_data <- data.frame(prison = 1, dose = 50, clinic = 1)
quant_p <- c(0.25, 0.5, 0.75)
pred_weimod <- predict(wei_mod, newdata = new_data, type = "quantile", p = quant_p)
pred_weimod
cbind(quant_p, pred_weimod)
```

### plot

```{r}
plot(x = predict(wei_mod, newdata = new_data, type = "quantile", p = (1 : 98) / 100), y = (1 : 98) / 100 , type = "b")
```

## Log-logistic model

```{r}
logl_mod <- survreg(Surv(survt, status == 1) ~ prison + dose + clinic, data = addicts1, dist = "loglogistic")
summary(logl_mod)
```

## PH assumption

```{r}
kmfit <- survfit(Surv(survt, status == 1) ~ clinic, data = addicts1)
plot(kmfit)
plot(log(kmfit$time), log(kmfit$surv/(1-kmfit$surv)))
```

## Survival Probability Plot

```{r}
lung <- lung #lung dataset from survival package
lung
```

```{r}
weimod_lung <- survreg(Surv(time, status == 2) ~ ph.ecog, data = lung, dist = "weibull")
summary(weimod_lung)
```


```{r}
pct <- 1:98/100
newdata_lung <- data.frame(ph.ecog = 2)
ptime <- predict(weimod_lung, newdata = newdata_lung, type = "quantile", p = pct, se = T)

matplot(cbind(ptime$fit, ptime$fit + 1.96*ptime$se.fit,
              ptime$fit - 1.96*ptime$se.fit)/30.5, 1-pct,
        xlab = "Months", ylab = "Survival", type = "l", lty = c(1,2,2), col = 1)
```


